Towards open collaboration in biomedical and health research: An open collaborative scoping review

### 2025-01-08 doc/paper.qmd
docs: :memo: expand on the introduction section of paper

-   Open science and its role in research

-   Collaboration and its role in modern research

    -   collaboration spanning from lab, to research center to multiple
        centers
    -   collaboration needed to tackle limitations: small populations,
        lack of replication, etc
    -   re-inventing discovery: a new way of working in science

Scientific research now almost always requires working with other
people.

With the growing complexity and specialization in scientific practices
and methods, together with globalisation of health and environmental
issues, there is a great need for a paradigm shift in research
collaboration to be able to tackle these challenges.

-   Open collaboration - a combination of two themes

    -   definition
    -   big potential in science
    -   but there very few resources and examples of how to integrate
        open science into collaborations

We define open collaboration using the definition as found in
[@Forte2013]:

@@

With the increasing emphasis on and demand for science to be more open,
how we collaborate together is a key component to making science more
open from the start of any project. But how do we collaborate in an open
and transparent way? What are the best practices and tools we can use?
What is an ideal collaborative workflow and how close or far are we from
this ideal in reality?

@@

-   Provide an overview of current practices of or opinions about
    research collaboration that follow basic open principles (e.g.,
    transparency, accessibility)

-   Summarize existing online tools and resources available to improve
    open collaboration in research

>>>>

The time when a single scientist could work alone in a lab and make
meaningful progress in their field is long gone. Scientific research in
the modern era, especially in certain fields, almost requires working
with other people, oftentimes a large number of people. Research
questions are increasingly more challenging to answer as they require
more time, resources, expertise, and personnel in order to meaningfully
answer.

With the current needs and limitations such as resources when doing
scientific research, collaborations range from those within a single
group or lab, to a few researchers across different departments in the
same institution, to many researchers and personnel spread across
multiple research centers. Collaborations are needed to address basic
limitations of working with small or constrained populations, to sharing
valuable but expensive equipment, or to assisting with technical
expertise or domain knowledge.

At the same time, there is also a higher need for science to be more
open, higher quality, and more rigorous[@Cole2024, @Umbach2024], which
necessitates that researchers collaborate with more researchers and with
a more diverse group of researchers to fulfill these needs. Open
scientific practices, such as open data, open source, and open
materials, aim to improve the reliability and accessibility of
scientific output to increase the societal impact of the results.

There are many hypothetical benefits to openness in science, including
in collaboration. For instance:

-   Greater transparency and accountability, because the built-in
    openness makes it easy to see what was done, how, and why.
-   Better reproducibility and inspectability, because everything can be
    easily viewed and can be verified, which is a key tenet of science.

The intersection between greater collaboration and more openness
highlights key challenges researchers face. How can we effectively
collaborate together in a way that fulfills the demand for more
openness, transparency, and accountability, while at the same time also
not substantially impacting our productivity (or ideally improving it).
How can we collaborate in an open and transparent way? What are the best
practices and tools we can use? What is an ideal collaborative workflow
and how close or how far are we from this ideal in reality?

From our own personal experience working in the field of health
research, we have varied and diverse ways of collaborating. Often the
most commonly used approach to collaboration is emailing Word documents
around to get feedback and using meetings to discuss and agree on
things. Rarely do we systematically examine whether we are following
best practices for collaboration nor do we collaborate in an open way
following the principles of open science. This lack of a clear standard
or approach to collaborating effectively and openly can greatly impact
productivity and quality of scientific research. It emphasizes a greater
need for a paradigm shift in how we researchers collaborate to be able
to effectively and efficiently tackle the challenges of the modern era.

For this scoping review, we define open collaboration using the
definition as found in [@Forte2013]:

@@

Open collaboration has a lot of potential benefits, many of which are
the same benefits as seen with open science. For instance, open
collaboration can lead to:

-   Easier onboarding of new collaborators, since everything is open and
    easily findable.
-   Higher collaboration opportunities, because more openness means more
    visibility as more parts of the scientific workflow are
    disseminated.
-   Faster time to dissemination, because the built-in openness makes it
    easier to disseminate results.

Unfortunately, there are, to our knowledge, very few resources and
examples of how to integrate open science into collaborations and how to
make collaboration be open.

@@

1.  Provide an overview of current practices of or opinions about
    research collaboration that follow basic open principles (e.g.,
    transparency, accessibility)

2.  Summarize existing online tools and resources available to improve
    open collaboration in research
